### 2023-03-20 3 Analysis Scripts/StatiscalAnalysis_ModelFitting.qmd

---
title: "StatiscalAnalysis_ModelFitting"
format: html
editor: visual
---

# Ciconia ciconia analysis script

#### this script loads the processed, cleaned data, does a simple analysis and saves the results to the results folder

```{r}
#load needed packages. make sure they are installed.
library(ggplot2) #for plotting
library(ggeffects) #to explore plots  
library(broom) #for cleaning up output from lm()
library(broom.mixed) #for cleaning up output from lmer()
library(here) #for data loading/saving
library(multilevelmod) #mixed effect models
library(tidymodels)
library(lme4)
library(sjPlot) #to help plot models
library(performance) #evaluate model fit and performance

#path to data
data_location <- here::here("2 Clean Data","stork_AMR_2021.rds")

#load data. 
data <- readRDS(data_location)
```

## Data fitting/statistical analysis

Data of interest: Landfill use index (LUI) and multi-drug resistence (MDR) throughout the different sampling periods during the stork breeding season.

We'll start by fitting two models. We will used mixed-effects models to control for repeated sampling of the same nest. The design is unbalanced, thus we cannot use a repeated measures anova.

We'll start with a single factor repeated measures evaluation the effect of LUI on blaTEM (betalactam) resistence gene burden.

```{r}
data(data, package = "lme4")

data %>% 
  ggplot(aes(x = samp, y = Burden)) + 
  geom_point() + 
  geom_line() +
  facet_wrap(~ nes) 
```

No obvious trends...

Build models using lmer and glmer

```{r}
lmer_spec <- 
  linear_reg() %>% 
  set_engine("lmer")

glmer_spec <- 
  linear_reg() %>% 
  set_engine("glmer")

```

Multi-drug resistance (0,1 data) vs All

```{r}
lmer_fit_null <- 
  lmer_spec %>% 
  fit(MDR ~ (1|nes), data = data)

lmer_fit_global <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui + s.age + samp + s.nsuccess +(1|nes), data = data)

#Univariate models
lmer_fit_1 <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui +(1|nes), data = data)

lmer_fit_2 <- 
  lmer_spec %>% 
  fit(MDR ~ s.age  +(1|nes), data = data)

lmer_fit_3 <- 
  lmer_spec %>% 
  fit(MDR ~ samp +(1|nes), data = data)

lmer_fit_4 <- 
  lmer_spec %>% 
  fit(MDR ~ s.nsuccess +(1|nes), data = data)

lmer_fit_5 <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui + samp +(1|nes), data = data)

lmer_fit_6 <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui + s.nsuccess +(1|nes), data = data)

lmer_fit_7 <- 
  lmer_spec %>% 
  fit(MDR ~ samp + s.nsuccess +(1|nes), data = data)

lmer_fit_8 <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui*s.age + samp + s.nsuccess +(1|nes), data = data)

lmer_fit_9 <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui + s.age*samp + s.nsuccess +(1|nes), data = data)

lmer_fit_10 <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui + s.age + samp*s.nsuccess +(1|nes), data = data)

lmer_fit_11 <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui + s.age +(1|nes), data = data)

lmer_fit_12 <- 
  lmer_spec %>% 
  fit(MDR ~ s.lui + s.age +  s.nsuccess +(1|nes), data = data)
```

Use AIC to see the best model; None look great...

```{r}
AIC_table1 <- compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12)

print(AIC_table1)
AIC_table1 = here("4 Results","Tables", "ModelAIC1.rds")
saveRDS(AIC_table1, file = AIC_table1)

plot(compare_performance(lmer_fit_null,lmer_fit_global, lmer_fit_1, lmer_fit_2, lmer_fit_3, lmer_fit_4, lmer_fit_5, lmer_fit_6, lmer_fit_7, lmer_fit_8, lmer_fit_9, lmer_fit_10, lmer_fit_11, lmer_fit_12, rank = TRUE))
```

Summarize and graph best model (lowwest AIC)

```{r}
check_model(lmer_fit_11)

# place results from fit into a data frame with the tidy function

model1 <- broom::tidy(lmer_fit_11)
print(model1)

# save fit results table  
model1 = here("4 Results", "Models", "LMERmodel1.rds")
saveRDS(model1, file = model1)

MLp <- plot_model(lmer_fit_11,terms = 's.lui', type="pred", 
           show.data = TRUE, show.values = TRUE, show.p = TRUE,
           title = "", show.legend = TRUE, line.size=1,
           axis.title = c("Landfill Use Index",'Multi-Drug Resistence'))

MLp

ggsave(MLp, file="LUIvsMDR_LME.tiff")
```

AMR Gene Burden vs All

```{r}
lmer_fit2_null <- #singular
  lmer_spec %>% 
  fit(Burden ~ (1|nes), data = data)

lmer_fit2_global <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui + s.age + samp + s.nsuccess +(1|nes), data = data) #overfit

#Univariate models
lmer_fit2_1 <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui +(1|nes), data = data)

lmer_fit2_2 <- 
  lmer_spec %>% 
  fit(Burden ~ s.age  +(1|nes), data = data) #overfit

lmer_fit2_3 <- 
  lmer_spec %>% 
  fit(Burden ~ samp +(1|nes), data = data)

lmer_fit2_4 <- 
  lmer_spec %>% 
  fit(Burden ~ s.nsuccess +(1|nes), data = data)

lmer_fit2_5 <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui + samp +(1|nes), data = data)

lmer_fit2_6 <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui + s.nsuccess +(1|nes), data = data)

lmer_fit2_7 <- 
  lmer_spec %>% 
  fit(Burden ~ samp + s.nsuccess +(1|nes), data = data)

lmer_fit2_8 <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui*s.age + samp + s.nsuccess +(1|nes), data = data) #overfit

lmer_fit2_9 <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui + s.age*samp + s.nsuccess +(1|nes), data = data) #overfit

lmer_fit2_10 <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui + s.age + samp*s.nsuccess +(1|nes), data = data) #overfit

lmer_fit2_11 <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui + s.age + s.nsuccess +(1|nes), data = data) #overfit

lmer_fit2_12 <- 
  lmer_spec %>% 
  fit(Burden ~ s.lui + s.age +(1|nes), data = data) #overfit

#Standardizing variables helped with the overfit/singular model results
```

Use AIC to see the best model; This doesnt look great either...

```{r}
AIC_table2 <- compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12)

print(AIC_table2)
AIC_table2 = here("4 Results","Tables", "ModelAIC2.rds")
saveRDS(AIC_table2, file = AIC_table2)

plot(compare_performance(lmer_fit2_null,lmer_fit2_global, lmer_fit2_1, lmer_fit2_2, lmer_fit2_3, lmer_fit2_4, lmer_fit2_5, lmer_fit2_6, lmer_fit2_7, lmer_fit2_8, lmer_fit2_9, lmer_fit2_10, lmer_fit2_11, lmer_fit2_12, rank = TRUE))
```

Summarize and graph best model (lowwest AIC)

```{r}
check_model(lmer_fit2_5)

model2 <- broom::tidy(lmer_fit2_5)
print(model2)

# save fit results table  
model2 = here("4 Results","Models", "LMERmodel2.rds")
saveRDS(model2, file = model2)

BLp <- plot_model(lmer_fit2_5,terms = 's.lui', type="pred", 
           show.data = TRUE, show.values = TRUE, show.p = TRUE,
           title = "", show.legend = TRUE, line.size=1,
           axis.title = c("Landfill Use Index",'Total Antibiotic Resistant Gene Burden'))

BLp

ggsave(BLp, file="LUIvsBurden_LME.tiff")
```

Preliminary results

Questions: 
Should over fit models be taken out of performance analysis?
What is the best tool/package to control for small data sets?