### 2023-04-26 3 Analysis Scripts/crossvalidation.qmd

---
title: "Cross Validation"
format: html
editor: visual
---

# Ciconia ciconia analysis script

#### this script loads the processed, cleaned data, does a simple analysis and saves the results to the results folder

```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(ggplot2) #for plotting
library(ggeffects) #to explore plots  
library(broom) #for cleaning up output from lm()
library(broom.mixed) #for cleaning up output from lmer()
library(here) #for data loading/saving
library(multilevelmod) #mixed effect models
library(tidymodels)
library(lme4)
library(sjPlot) #to help plot models
library(performance) #evaluate model fit and performance
library(see)
library(cAIC4)
library(rpart)
library(rpart.plot)
library(glmertree)

#path to data
data_location <- here::here("2 Clean Data","stork_AMR_clean.rds")

#load data. 
data <- readRDS(data_location)
```

```{r}
# Set seed
set.seed(2023)

# Put 3/4 of the data into the training set 
data_split <- initial_split(data, prop = 3/4)

# Create data frames for the two sets
train_data <- training(data_split)
test_data  <- testing(data_split)
```

```{r}
# 5-fold cross-validation, 5 times repeated, using MDR as strata 
folds <- vfold_cv(train_data, v = 5, repeats = 5, strata = MDR)

#create model and set engine
tree_model <- 
  decision_tree(
    cost_complexity = tune(),
    tree_depth = tune()) %>% 
  set_engine("rpart") %>% 
  set_mode("regression")

#create tree workflow for global model
tree_workflow_global <- workflow() %>%
  add_model(tree_model) %>%
  add_formula(Burden ~ s.lui + s.age + samp + s.nsuccess + nes)

#tree grid
tree_grid <- grid_regular(cost_complexity(),
                          tree_depth(),
                          levels = 5)

#tree_resamples
tree_resamp <- 
  tree_workflow_global %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_resamp %>%
  collect_metrics()

tree_resamp %>%
  autoplot()
```
```{r}
#create tree workflow for null
tree_workflow_null <- workflow() %>%
  add_model(tree_model) %>%
  add_formula(Burden ~ 1)

#tree_resamples
tree_resamp_null <- 
  tree_workflow_null %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_resamp_null %>%
  collect_metrics()

tree_resamp_null %>%
  autoplot()
```