### 2023-05-05 5 Manuscript/manuscript/supplement/Supplementary_Material.qmd
Final edits to supplement

---
title: "Supplementary Material"
subtitle: "A Match Made in Landfills? Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (Ciconia ciconia) throughout the breeding season in Madrid, Spain"
author: Raquel Francisco and Seth Lattner
date: "`r Sys.Date()`"
format: html
editor: visual
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Overview

Contained in this document is supplementary material for the manuscript, **"A Match Made in Landfills? Exploring the diversity and burden of antimicrobial resistance genes carried by white stork (*Ciconia ciconia*) throughout the breeding season in Madrid, Spain"**. Supplementary materials include instructions for reproducing data analysis and additonal model results.

# Code and file information

To reproduce the data analysis for this manuscript, raw data should first be cleaned by running the DataCleaningScript.qmd file in the "1 Data Cleaning Script" folder. Next, the Data_Exploratory_Analysis.qmd file from the "3 Analysis Scripts" should be run, followed by the StatisticalAnalysis_ModelFitting.qmd file from the same folder. To recreate the manuscript, the Manuscript.qmd file under the "5 Manuscript/manuscript" folders should be run. Raw data is housed in the file StorkAMRraw_2020_2021.csv within the "1 Data Cleaning Script/raw_data" folders. Processed data is housed in the stork_AMR_clean.rds file within the "2 Clean Data" folder. Additionally, a README.md file that serves as a codebook for this project can be found within the "1 Data Cleaning Script/raw_data" folders.

{{< pagebreak >}}

# Additional Method Details

Binomial-family generalized linear mixed models (GLMMs) and linear mixed models (LMMs) were used to assess the effects of multiple predictors on the presence of multi-drug resistance (MDR) and total antimicrobial resistance gene burden (AMR), respectively.

Additionally, a cross-validation approach using tree model was implemented. Prior to machine learning analysis, the data were split into training (75%) and testing (25%) groups. A 5-fold cross-validation, 5 times repeated approach was implemented using MDR as the strata. A tuning grid was utilized using 5 levels of tree depth. After fitting the tree model to the training data, the best model was selected using the area under the receiver operating characteristic curve (ROC AUC). This final model was then fit to the testing data.

{{< pagebreak >}}

# Additional results

Contained below are additional results from the full analyses.

## Full Analysis Results

A total of 14 GLMMs (including a global and null model) were run to assess the effects of several predictors on MDR. Based on Akaike's Information Criterion adjusted for small sample size (AIC*c*), model 11 (containing the predictors LUI and age) was the highest performing model. A total of 15 LMMs were run to assess the effects of several predictors on AMR. Based on AIC*c*, the global model was the highest performing model, followed by a univariate LMM containing age as the only predictor.

![Figure 1. Facet Plot of single factor repeated measures evaluation the effect of LUI on blaTEM (betalactam) resistence gene burden](Supplementary_Material_files/facet_plot.png)

### GLMM Model Results

![Figure 2. Comparison of Generalized Linear Mixed Model Indices](Supplementary_Material_files/glmer_indices.png)

![Figure 3. Model Check Metrics for Best-performing GLMM](Supplementary_Material_files/glmer_check.png)

### LMM Model Results

![Figure 4. Comparison of Linear Mixed Model Indices](Supplementary_Material_files/lmer_indices.png) ![Figure 5. Model Check Metrics for Best-performing LMM](Supplementary_Material_files/lmer_check.png)

### Tree Model Results

Cost-complexity parameters for various tree depths are shown in Figure 6.

![Figure 6. Cost-complexity Parameter Plot](Supplementary_Material_files/autoplot.png) Results from the finalized tree model can be found in Figure 7.

![Figure 7. Model Check Metrics for Best-performing LMM](Supplementary_Material_files/rpartplot2.png)

{{< pagebreak >}}

# Discussion

The tree model was ultimately not incorporated into the final manuscript. The ability to account for random effects using the GLMMs and LMMs was better suited for the questions being addressed and the data on hand.

{{< pagebreak >}}

# References